---
title: "convert DerM22 signature matrix from probe IDs to gene symbols"
subtitle: ""
date: "`r Sys.Date()`"
output:
  html_notebook:
    theme: readable
    code_folding: none
---

"Characterization of disease-specific cellular abundance profiles of chronic inflammatory skin conditions from deconvolution of biopsy samples" (https://doi.org/10.1186/s12920-019-0567-7)

```{r message=TRUE, warning=FALSE}
suppressPackageStartupMessages({
  library(tidyverse)
  library(glue)
})
```

import the original DerM22 matrix

```{r message=FALSE, warning=FALSE}
derm22_tbl = read_tsv("CIBERSORT.DerM22.probes.txt", show_col_types = FALSE)
dim(derm22_tbl)
```

```{r}
head(derm22_tbl)
```

```{r}
derm22_mat = derm22_tbl %>% column_to_rownames("probe_id") %>% as.matrix()
derm22_mat[1:5, 1:3]
```

import the original LM22 matrix

```{r message=FALSE, warning=FALSE}
lm22_tbl = read_tsv("CIBERSORT.LM22.txt", show_col_types = FALSE)
dim(lm22_tbl)
```

```{r}
lm22_mat = lm22_tbl %>% column_to_rownames("Gene symbol") %>% as.matrix()
lm22_mat[1:5, 1:3]
```

check the distribution of values

```{r}
hist(log(derm22_mat + 0.001), breaks = 50, col = "black")
```

```{r}
hist(log(lm22_mat + 0.001), breaks = 50, col = "black")
```

add gene names to probes

```{r}
mart = biomaRt::useMart("ensembl", dataset = "hsapiens_gene_ensembl", host = "https://may2021.archive.ensembl.org")
```

```{r}
probes_tbl = biomaRt::getBM(
  mart = mart,
  attributes = c("affy_hg_u133_plus_2", "external_gene_name"),
  filters = "affy_hg_u133_plus_2",
  values = rownames(derm22_mat),
  uniqueRows = TRUE
)
dim(probes_tbl)
```

```{r}
probes_tbl =
  probes_tbl %>%
  rename(probe_id = affy_hg_u133_plus_2, gene = external_gene_name)
head(probes_tbl)
```

remove probes without gene names

```{r}
probes_tbl = probes_tbl %>% filter(gene != "") %>% arrange(gene, probe_id)
dim(probes_tbl)
```

add variance to probes

```{r}
probes_tbl =
  probes_tbl %>%
  inner_join(
    tibble(
      probe_id = rownames(derm22_mat),
      variance = matrixStats::rowVars(derm22_mat)
    ),
    by = "probe_id"
  )
dim(probes_tbl)
```

for genes with multiple probes, select the probe with highest variance

```{r}
probes_tbl =
  probes_tbl %>%
  group_by(gene) %>%
  slice_max(variance, n = 1) %>%
  ungroup()
dim(probes_tbl)
```

replace probe IDs with gene names in the original table

```{r}
derm22_tbl =
  probes_tbl %>%
  select(probe_id, gene) %>%
  inner_join(derm22_tbl, by = "probe_id") %>%
  select(!probe_id)
dim(derm22_tbl)
```

```{r}
head(derm22_tbl)
```

save

```{r}
write_tsv(derm22_tbl, "CIBERSORT.DerM22.genes.txt")
```

session info

```{r}
sessionInfo()
```


