---
title: "CIBERSORT cell type deconvolution"
subtitle: "ST"
date: "`r Sys.Date()`"
output:
  html_notebook:
    theme: readable
    code_folding: none
---

```{r message=TRUE, warning=FALSE}
suppressPackageStartupMessages({
  library(tidyverse)
  library(glue)
  library(pheatmap)
  library(ggplot2)
  library(cowplot)
  library(ggsci)
})
```

```{r}
source("./CIBERSORT.R")
```

# define inputs/outputs

input

```{r}
subset_label = "healthy"
subset_label = "all"
subset_label = "all-per-group"

exp_csv = glue("./data/{subset_label}.expression.mean.csv")
if (!file.exists(exp_csv)) stop()
```

output

```{r}
out_dir = "./out-celltypes-cibersort"
dir.create(out_dir, showWarnings = FALSE)
out_dir = glue("{out_dir}/{subset_label}")
dir.create(out_dir, showWarnings = FALSE)
```

CIBERSORT input: "should be in non-log linear space with positive numerical values"

```{r message=FALSE, warning=FALSE}
exp_tbl = read_csv(exp_csv, show_col_types = FALSE)
dim(exp_tbl)
```

```{r}
head(exp_tbl)
```

check the distribution of values for one sample

```{r}
example_exp = exp_tbl %>% pull(2)
example_exp = example_exp[example_exp > quantile(example_exp, 0.05) & example_exp < quantile(example_exp, 0.95)]
hist(example_exp, breaks = 50, col = "black")
```

check the total counts

```{r}
exp_tbl %>% select(!gene) %>% colSums() %>% quantile() %>% scales::comma()
```

save expression table as a tab-delimited file

```{r}
mixture_txt = glue("{out_dir}/cibersort.mixture.tsv")
write_tsv(exp_tbl, mixture_txt)
```

run CIBERSORT analysis

"we recommend that RNA-Seq data not undergo quantile normalization (enabled by default)"

```{r message=FALSE, warning=FALSE}
signature_txt = "CIBERSORT.LM22.txt"
cibersort_mat =
  CIBERSORT(
    sig_matrix = signature_txt,
    mixture_file = mixture_txt,
    perm = 1000,
    QN = FALSE
  )
```

clean up and save the CIBERSORT matrix

```{r}
cibersort_tbl =
  t(cibersort_mat) %>%
  round(5) %>%
  as.data.frame() %>%
  rownames_to_column("cell_type") %>%
  as_tibble() %>%
  filter(!cell_type %in% c("P-value", "Correlation", "RMSE"))
write_csv(cibersort_tbl, glue("{out_dir}/celltypes-cibersort.csv"))
cibersort_tbl
```

filter out cell types with only low scores

```{r}
cibersort_mat_filtered =
  cibersort_tbl %>% as.data.frame() %>% column_to_rownames("cell_type") %>% as.matrix()
cibersort_mat_filtered[1:10, 1:5]
```

heatmap

```{r}
heatmap_colors = colorRampPalette(c("#FFFFFF", "#E41A1C"))(50)
pheatmap(
  cibersort_mat_filtered, border_color = NA, color = heatmap_colors,
)
pheatmap(
  cibersort_mat_filtered, border_color = NA, color = heatmap_colors,
  fontsize_row = 12, fontsize_col = 12,
  filename = glue("{out_dir}/celltypes-cibersort.heatmap.png"), width = 10, height = 8
)
pheatmap(
  cibersort_mat_filtered, border_color = NA, color = heatmap_colors,
  fontsize_row = 12, fontsize_col = 12,
  filename = glue("{out_dir}/celltypes-cibersort.heatmap.pdf"), width = 10, height = 8
)
```

```{r}
cibersort_tbl_long = cibersort_tbl %>% gather(key = "sample", value = "fraction", -cell_type)

# select population for ordering samples
sample_order =
  cibersort_tbl_long %>%
  filter(cell_type == "Macrophages M2") %>%
  arrange(-fraction) %>%
  pull(sample)

bar_plot = ggplot(cibersort_tbl_long) +
  geom_bar(
    aes(
      x = fct_relevel(sample, sample_order),
      y = fraction,
      fill = fct_reorder(cell_type, fraction, .desc = TRUE)
    ),
    stat = "identity") +
  scale_y_continuous(expand = c(0, 0)) +
  theme_cowplot() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  labs(x = "sample", fill = "cell type") +
  scale_fill_igv()
save_plot(glue("{out_dir}/celltypes-cibersort.barplot.png"), plot = bar_plot, base_height = 6, base_width = 12)
save_plot(glue("{out_dir}/celltypes-cibersort.barplot.pdf"), plot = bar_plot, base_height = 6, base_width = 12)
bar_plot
```

session info

```{r}
sessionInfo()
```


