---
title: "Hannifa et al. single cell - healthy skin data subset"
output: github_document
---

# ST HEALTHY SAMPLES - PART 3

## Integration with REYNOLDS et al 

## (DOI - <https://doi.org/10.1126/science.aba6500>).

### Load all of the required packages

```{r}
library(tidyverse)
library(Seurat)
library(cowplot)
library(RColorBrewer)
library(pheatmap)
library(scales)
```

### Import functions from the custom R script

```{r}
source("SPATIAL_FUNCTIONS.R")
```

### Load processed single cell R data.

(*replace this with zenodo repo link for final release)*

```{r}
load("../../scRNA_DATA/HEALTHY_SC_RNA_PROCESSED.RData")
```

### UMAP plot for healthy skin data

### (Using final_clustering labels from the original manuscript)

```{r}
healthy_skin.sc_data <- FindNeighbors(healthy_skin.sc_data, dims = 1:40)
healthy_skin.sc_data <- FindClusters(healthy_skin.sc_data, verbose = FALSE,resolution = 1)
healthy_skin.sc_data <- RunUMAP(healthy_skin.sc_data, dims = 1:40)
```

### Remove unlabeled cells (NAN) from the dataset

```{r}
Idents(healthy_skin.sc_data) <- "final_clustering"
healthy_skin.sc_data <- subset(healthy_skin.sc_data,idents = c("nan"),invert=TRUE)
```

```{r}
healthy_skin.sc_data <- readRDS("../scRNA_DATA/healthy_skin.sc_data_HNF.RDS")
```

```{r}
#pdf(width = 12,height=8,file = "UMAP_HANNIFA_DATA_HEALTHY_SAMPLES_ONLY.pdf")
DimPlot(healthy_skin.sc_data,pt.size = 3.5,raster=FALSE)
#dev.off()
```

```{r}

Idents(healthy_skin.sc_data) <- "final_clustering"
DimPlot(healthy_skin.sc_data,raster = FALSE)

pdf(file = "UMAP_ONLY_HEALTHY_SAMPLES_(HANNIFA_PAPER).pdf",width = 16,height=8)
print(DimPlot(healthy_skin.sc_data_subset,pt.size = 1.2,shuffle = TRUE,raster=FALSE))
dev.off()
```

### Identify marker genes for the final_clustering labels

```{r,eval=FALSE}
Idents(healthy_skin.sc_data) <- "final_clustering"
hnf_skin_scRNA.markers <- FindAllMarkers(healthy_skin.sc_data,assay = "RNA",logfc.threshold = 0.25,max.cells.per.ident=400,only.pos = TRUE)
write.csv(hnf_skin_scRNA.markers,"../../scRNA_DATA/HANNIFA_MARKERS.csv")
```

### Load marker genes from the saved file for later use

```{r}
hnf_skin_scRNA.markers <- read.csv("../../scRNA_DATA/HANNIFA_MARKERS.csv") %>% filter(p_val_adj<=0.05) %>% group_by(cluster) %>% top_n(n =300,wt = avg_log2FC) %>% filter(avg_log2FC>0.25)
```

### Cell counts / frequency for each cell type in the processed Seurat object

```{r}
hnf_data_counts <- as.data.frame(table(healthy_skin.sc_data@meta.data$final_clustering))
write.csv(hnf_data_counts,file="HEALTHY_SAMPLES_HANNIFA_DATA_COUNTS.csv")
```

### MIA ( Multi-modal intersection analysis)

### original publication- (<https://doi.org/10.1038/s41587-019-0392-8>)

#### pre-computation parameters

#### Number of intersecting / common genes between ST and scRNA data -set as background genes.

```{r}
##INTERSECT GENES BETWEEN scRNA and Spatial data
st.genes <- unique(rownames(new.skin.combined@assays$Spatial@counts))
sc.genes <- unique(rownames(healthy_skin.sc_data@assays$RNA@counts))
all.genes.scrna_and_spt <- unique(intersect(sc.genes,st.genes))
```

### Calculate cell type enrichment scores for all cell types in the dataset - healthy sample data only

```{r}
MIA_results_PART_2 <- MIA(total_genes = length(all.genes.scrna_and_spt),single_cell.markers = hnf_skin_scRNA.markers,spatial.markers = filtered_spatial_markers)

E.data_PART_2 <- MIA_results_PART_2 %>% column_to_rownames("cluster")
E.data_PART_2 <- E.data_PART_2[,order(colnames(E.data_PART_2))]
is.na(E.data_PART_2) <- do.call(cbind,lapply(E.data_PART_2, is.infinite))
pheatmap(E.data_PART_2,cluster_cols = FALSE,cluster_rows = FALSE,fontsize=15,color = col.pal)
```

### Immune cell types only

```{r}
## IMMUNE CELLS
immune_only.E.data_PART_2 <- E.data_PART_2[,c("DC1","DC2","Macro_1","Macro_2","moDC_1","moDC_2","moDC_3","MigDC","Mono","Inf_mono","ILC1_NK","ILC1_3","ILC2","Tc","Th","Treg","Mast_cell","Plasma","NK")]
pdf(file = "MIA_regions_(HANNIFA_PAPER)_IMMUNE_CELLS.pdf",width = 10,height = 5)
print(pheatmap(immune_only.E.data_PART_2,cluster_cols = FALSE,cluster_rows = FALSE,fontsize=15,color = col.pal))
dev.off()

```

### Non-Immune / Structural cell types

```{r}
non_immune_only.E.data_PART_2 <- E.data_PART_2 %>% dplyr::select(-c("DC1","DC2","Macro_1","Macro_2","moDC_1","moDC_2","moDC_3","MigDC","Mono","Inf_mono","ILC1_NK","ILC1_3","ILC2","Tc","Th","Treg","Mast_cell","Plasma","NK","nan"))
pdf(file = "MIA_regions_(HANNIFA_PAPER)_NON_IMMUNE_CELLS.pdf",width = 10,height = 5)
print(pheatmap(non_immune_only.E.data_PART_2,cluster_cols = FALSE,cluster_rows = FALSE,fontsize=15,color = col.pal))
dev.off()
```

## 
