---
title: "Sullivan et al. single cell data - Preadipocytes and Adipose precursor cells"
output: github_document
---

# ST HEALTHY SAMPLES - PART 4

## Integration with Sullivan et al

## (DOI- <https://doi.org/10.1038/s41590-021-00922-4>)

```{r}
library(Seurat)
library(tidyverse)
library(extrafont) 
#font_import()
#loadfonts(device = "win")
```

### Loading data from Sullivan et al

Only importing healthy / lean samples from the paper.

```{r,eval=FALSE}
lean_1.cd45_neg <- data.table::fread("https://ftp.ncbi.nlm.nih.gov/geo/samples/GSM4717nnn/GSM4717152/suppl/GSM4717152_CD45N-L1.csv.gz")
lean_2.cd45_neg <- data.table::fread("https://ftp.ncbi.nlm.nih.gov/geo/samples/GSM4717nnn/GSM4717153/suppl/GSM4717153_CD45N-L2.csv.gz")
lean_3.cd45_neg <- data.table::fread("https://ftp.ncbi.nlm.nih.gov/geo/samples/GSM4717nnn/GSM4717154/suppl/GSM4717154_CD45N-L3.csv.gz")
lean.DC <- data.table::fread("https://ftp.ncbi.nlm.nih.gov/geo/samples/GSM4724nnn/GSM4724865/suppl/GSM4724865_DC_Lean.csv.gz")
lean.MAC <- data.table::fread("https://ftp.ncbi.nlm.nih.gov/geo/samples/GSM4724nnn/GSM4724867/suppl/GSM4724867_Mac_Lean.csv.gz")
lean.cd200r_neg <- data.table::fread("https://ftp.ncbi.nlm.nih.gov/geo/samples/GSM4724nnn/GSM4724861/suppl/GSM4724861_CD200r_Negative_Lean.csv.gz")
lean.cd200r_pos <- data.table::fread("https://ftp.ncbi.nlm.nih.gov/geo/samples/GSM4724nnn/GSM4724863/suppl/GSM4724863_CD200r_Positive_Lean.csv.gz")
```

### Add gene names / HGNC symbols to the raw matrix

```{r,eval=FALSE}
library(biomaRt)
ensembl <- useEnsembl(biomart = "genes", dataset = "hsapiens_gene_ensembl")
attributes = listAttributes(ensembl)
Convert.ids <- lean_1.cd45_neg$V1
Gene.names <- getBM(attributes = c('hgnc_symbol','ensembl_gene_id'),
                    filters = 'ensembl_gene_id',
      values = Convert.ids, 
      mart = ensembl)
lean_1.cd45_neg <- lean_1.cd45_neg %>% as_tibble()
Gene.names <- Gene.names %>% filter (! duplicated('ensembl_gene_id')) %>% as_tibble()

```

```{r,eval=FALSE}
lean_1.cd45_neg <- lean_1.cd45_neg[lean_1.cd45_neg$V1 %in% Gene.names$ensembl_gene_id ,]
lean_2.cd45_neg <- lean_2.cd45_neg[lean_2.cd45_neg$V1 %in% Gene.names$ensembl_gene_id ,] 
lean_3.cd45_neg <-lean_3.cd45_neg[lean_3.cd45_neg$V1 %in% Gene.names$ensembl_gene_id ,] 
lean.DC <- lean.DC[lean.DC$V1 %in% Gene.names$ensembl_gene_id ,] 
lean.MAC <- lean.MAC[lean.MAC$V1 %in% Gene.names$ensembl_gene_id ,] 
lean.cd200r_neg <-lean.cd200r_neg[lean.cd200r_neg$V1 %in% Gene.names$ensembl_gene_id ,] 
lean.cd200r_pos <- lean.cd200r_pos[lean.cd200r_pos$V1 %in% Gene.names$ensembl_gene_id ,] 

```

```{r,eval=FALSE}
lean_1.cd45_neg_with_symbols <- left_join(lean_1.cd45_neg,Gene.names,by=c("V1"="ensembl_gene_id"),keep=TRUE) %>% filter(!is.na("hgnc_symbol") | "hgnc_symbol" != "") %>%  distinct(ensembl_gene_id, .keep_all = TRUE) %>% dplyr::select(-c("V1","ensembl_gene_id")) %>%  distinct(hgnc_symbol, .keep_all = TRUE) %>% column_to_rownames("hgnc_symbol") %>% filter(rowMeans(is.na(.)) < 0.1)

other.samples <- list(lean_2.cd45_neg,lean_3.cd45_neg,lean.DC,lean.MAC,lean.cd200r_neg,lean.cd200r_pos)

```

```{r,eval=FALSE}
i <- 1
while(i <= length(other.samples)){
  filtered_data <- left_join(other.samples[[i]],Gene.names,by=c("V1"="ensembl_gene_id"),keep=TRUE) %>% filter(!is.na("hgnc_symbol") | "hgnc_symbol" != "") %>%  distinct(ensembl_gene_id, .keep_all = TRUE) %>% dplyr::select(-c("V1","ensembl_gene_id")) %>%  distinct(hgnc_symbol, .keep_all = TRUE) %>% column_to_rownames("hgnc_symbol") %>% filter(rowMeans(is.na(.)) < 0.1)
  other.samples[[i]] <- filtered_data
  i <- i+1
}
```

### Generate seurat objects from scRNA matrix.

```{r,eval=FALSE}
lean_1.cd45_neg.obj <- CreateSeuratObject(counts=lean_1.cd45_neg_with_symbols,project = "LEAN DONOR 1 - CD45 -")
lean_2.cd45_neg.obj <- CreateSeuratObject(counts=other.samples[[1]],project = "LEAN DONOR 2 - CD45 -")
lean_3.cd45_neg.obj <- CreateSeuratObject(counts=other.samples[[2]],project = "LEAN DONOR 3 - CD45 -")
lean.DC.obj <- CreateSeuratObject(counts=other.samples[[3]],project = "LEAN DONOR - DC")
lean.MAC.obj <- CreateSeuratObject(counts=other.samples[[4]],project = "LEAN DONOR - MAC")
lean.cd200r_neg.obj <- CreateSeuratObject(counts=other.samples[[5]],project = "LEAN DONOR - CD200R -")
lean.cd200r_pos.obj <- CreateSeuratObject(counts=other.samples[[6]],project = "LEAN DONOR 3 - CD200R +")
```

### Combine all lean samples into one seurat object

```{r,eval=FALSE}
combined_lean_samples.obj <- merge(lean_1.cd45_neg.obj,c(lean_2.cd45_neg.obj,lean_3.cd45_neg.obj,lean.DC.obj ,lean.MAC.obj,lean.cd200r_neg.obj,lean.cd200r_pos.obj))
```

### QC step - calculate mt percent expression

```{r,eval=FALSE}
# calculate the percentage of mitochondrial genes here and store it in percent.mito using the AddMetaData
mt_genes = grep("^MT-", rownames(GetAssayData(combined_lean_samples.obj)), ignore.case = TRUE, value = TRUE)
percent_mt = Matrix::colSums(GetAssayData(combined_lean_samples.obj)[mt_genes, ]) / Matrix::colSums(GetAssayData(combined_lean_samples.obj))
percent_mt = round(percent_mt * 100, digits = 3)

combined_lean_samples.obj = AddMetaData(combined_lean_samples.obj, metadata = percent_mt, col.name = "percent.mt")
```

### General QC plot

```{r,eval=FALSE}
VlnPlot(combined_lean_samples.obj, features = c("nFeature_RNA", "nCount_RNA","percent.mt"), ncol = 3)
```

### Filter samples using same QC parameters as the original paper

```{r,eval=FALSE}
combined_lean_samples.obj <- subset(combined_lean_samples.obj, subset = nFeature_RNA > 500 & percent.mt < 25)
```

### Normalization and Scaling data

```{r,eval=FALSE}
combined_lean_samples.obj <- NormalizeData(combined_lean_samples.obj, normalization.method = "LogNormalize", scale.factor = 10000)
combined_lean_samples.obj <- FindVariableFeatures(combined_lean_samples.obj, selection.method = "vst", nfeatures = 2000)
```

```{r,eval=FALSE}
all.genes <- rownames(combined_lean_samples.obj)
combined_lean_samples.obj <- ScaleData(combined_lean_samples.obj, features = all.genes)
```

### Run PCA

```{r,eval=FALSE}
combined_lean_samples.obj <- RunPCA(combined_lean_samples.obj, features = VariableFeatures(object = combined_lean_samples.obj))
```

### UMAP and clustering

```{r,eval=FALSE}
combined_lean_samples.obj <- FindNeighbors(combined_lean_samples.obj, dims = 1:20)
combined_lean_samples.obj <- FindClusters(combined_lean_samples.obj, resolution = 0.5)
```

```{r,eval=FALSE}
combined_lean_samples.obj <- RunUMAP(combined_lean_samples.obj, dims = 1:20)
```

### Save the output into RDS file

```{r,eval=FALSE}
saveRDS("../../scRNA ADIPOCYTES/LEAN_SAMPLES_SCRNA.RDS")
```

### Load the pre-processed data - RDS object

```{r}
combined_lean_samples.obj <- readRDS("../../scRNA ADIPOCYTES/LEAN_SAMPLES_SCRNA.RDS")
```

### Plot the UMAP (added cell type labels based on marker genes)

```{r}
Idents(combined_lean_samples.obj) <- "seurat_clusters"
DimPlot(combined_lean_samples.obj,label = TRUE)
```

```{r}
FeaturePlot(combined_lean_samples.obj,features = c("CD34","DPP4","APOD","F3"))

# ADIPOCYTES AND SMOOTH MUSCLE
FeaturePlot(combined_lean_samples.obj,features = c("TAGLN","HEXIM1"))

FeaturePlot(combined_lean_samples.obj,features = c("VWF","CDH5"))
```

# 

```{r,eval=FALSE}
saveRDS(combined_lean_samples.obj,file="../../scRNA ADIPOCYTES/LEAN_SAMPLES_SCRNA.RDS")
```

### Find Marker genes, plot a heatmap to label clusters by marker genes

```{r}
combined_lean_samples.MARKERS <- FindAllMarkers(combined_lean_samples.obj, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25,assay = "RNA")
top10 <- combined_lean_samples.MARKERS %>%
    group_by(cluster) %>%
    filter(p_val_adj<0.05) %>%
    top_n(n = 10, wt = avg_log2FC) 
DoHeatmap(combined_lean_samples.obj, features = top10$gene,assay = "RNA") + NoLegend()
```

### Label cell types of interest

```{r}
Idents(combined_lean_samples.obj) <- "seurat_clusters"
cluster.ids <- c("Preadipocytes","Preadipocytes","2","APCs","4","5","6","7","8","9","10","11","12","13","14","15","16","17","18")
names(cluster.ids) <- levels(combined_lean_samples.obj)
combined_lean_samples.obj <- RenameIdents(combined_lean_samples.obj, cluster.ids)
combined_lean_samples.obj$CellType <- Idents(combined_lean_samples.obj)
```

```{r}
DimPlot(combined_lean_samples.obj,label = TRUE)
```

### Load the Healthy ST dataset

```{r}
load("../../scRNA_DATA/HEALTHY_SC_RNA_PROCESSED.RData")
```

### Calculate enrichment for each cell type (Using Seurat Anchor Integration)

```{r,eval=FALSE}
# RUN THIS ON BIG PURPLE 
skin_reference.3 <- SCTransform(combined_lean_samples.obj, ncells = 3000, verbose = FALSE) %>% RunPCA(verbose = FALSE) %>% 
    RunUMAP(dims = 1:40)
anchors <- FindTransferAnchors(reference = skin_reference.3, query = new.skin.combined, normalization.method = "SCT")
predictions.assay <- TransferData(anchorset = anchors, refdata = skin_reference.3$CellType, prediction.assay = TRUE, 
    weight.reduction = new.skin.combined[["pca"]],dims = 1:40)
new.skin.combined[["predictions_lean_fat_data"]] <- predictions.assay
```

```{r}
##INTERSECT GENES BETWEEN scRNA and Spatial data
st.genes <- unique(rownames(new.skin.combined@assays$Spatial@counts))
sc.genes <- unique(rownames(combined_lean_samples.obj@assays$RNA@counts))
all.genes.scrna_and_spt <- unique(intersect(sc.genes,st.genes))
```

```{r}
library(RColorBrewer)
library(pheatmap)
col.pal <- RColorBrewer::brewer.pal(9, "OrRd")
```

```{r}
filtered_single_cell_markers <- combined_lean_samples.MARKERS %>% filter(p_val_adj<=0.05) %>% group_by(cluster) %>% top_n(n =300,wt = avg_log2FC) %>% filter(avg_log2FC>0.25)
filtered_spatial_markers <- new.skin.combined.markers %>% filter(p_val_adj<=0.05) %>% group_by(cluster) %>% top_n(n =300,wt = avg_log2FC) %>% filter(avg_log2FC>0.25)
MIA_results <- MIA(total_genes = length(all.genes.scrna_and_spt),single_cell.markers = filtered_single_cell_markers,spatial.markers = filtered_spatial_markers)

E.data <- MIA_results %>% column_to_rownames("cluster")
E.data <- E.data[,order(colnames(E.data))]
#pdf(file = "MIA_regions_SULLIVAN_DATA.pdf",width = 10,height = 4)
pheatmap(E.data,cluster_cols = FALSE,cluster_rows = FALSE,fontsize=15,color = col.pal)
#dev.off()
```

```{r}
#Ob (leptin), Hoxc8, and Hoxc9
#Mpzl2, Ebf3, and Fbox31
adipocyte_genes_1 <- c("FABP4","PLIN1","PLIN4","ADIRF","SCD","ADIPOQ","CAV1","LPL")
adipocyte_genes_2 <- c("LEP","HOXC8","HOXC9","MPZL2","EBF3","FBXO31")
VlnPlot(new.skin.combined,features = adipocyte_genes_1)
VlnPlot(new.skin.combined,features = adipocyte_genes_2)
```

# PATHWAYS USING ENRICH R

```{r}
library(enrichR)
setEnrichrSite("Enrichr")
```

```{r}
dbs <- listEnrichrDbs()
```

```{r}
dbs
websiteLive <- TRUE
```

```{r}
cluster.labels <- c("Preadipocytes","APC")
dbs <- c("Elsevier_Pathway_Collection", "MGI_Mammalian_Phenotype_Level_4_2021", "KEGG_2021_Human","GO_Biological_Process_2021","GO_Cellular_Component_2021","GO_Molecular_Function_2021")
dir.create("../../scRNA ADIPOCYTES/PATHWAYS-BY-CLUSTER/")
for(x in cluster.labels){
  de.genes <- FindMarkers(combined_lean_samples.obj,assay = "RNA",test.use = "wilcox",logfc.threshold = log2(1.1),ident.1 = x) %>% filter(p_val_adj<0.1) %>% rownames_to_column("Gene")
  up.genes <- de.genes %>% filter(avg_log2FC>0)
  down.genes <- de.genes %>% filter(avg_log2FC<0)
  if(!is.null(up.genes)){
    enriched <- enrichr(up.genes$Gene, dbs)
    for(a in dbs){
      write.csv(enriched[[a]],file=paste("../../scRNA ADIPOCYTES/PATHWAYS-BY-CLUSTER/",a,"CLUSTER_",x,"_UP_GENES.csv"),)
    }
  }
  if(!is.null(down.genes)){
    enriched <- enrichr(down.genes$Gene, dbs)
    for(b in dbs){
      write.csv(enriched[[b]],file=paste("../../scRNA ADIPOCYTES/PATHWAYS-BY-CLUSTER/",b,"CLUSTER_",x,"_DOWN_GENES.csv"),)
    }
  }

}
```
