---
title: "TRAVIS HUGHES et al single cell- healthy skin subset"
output: github_document
---

## ST HEALTHY SAMPLES
# Integeration with TRAVIS HUGHES et al.

Load all of the required packages
```{r}
library(tidyverse)
library(Seurat)
library(cowplot)
library(RColorBrewer)
library(pheatmap)
library(scales)
```

# Import functions from the custom R script
```{r}
source("SPATIAL_FUNCTIONS.R")
```

## LOAD ST HEALTHY SKIN DATA
```{r}
load(file = "HEALTHY_SKIN_SAMPLES_ST.RData")
```

## SINGLE CELL DATA FROM TRAVIS PAPER
```{r}
load("../../TRAVIS_scRNA_DATA/SkinSeuratTotal.Rdata")
```

```{r}
# META DATA FROM TRAVIS 
meta.data <- read.csv("../../TRAVIS_scRNA_DATA/Cell_Level_Metadata.csv")
```

# ccopy meta.data to the seurat object
```{r}
#AddMetaData(SkinSeuratTotal,metadata = meta.data$Specific_CellType,col.name = "Specific_Celltype")
skin.meta.data <- as.data.frame(SkinSeuratTotal@meta.data) %>% rownames_to_column("CellID")
inner_join(x=skin.meta.data,y=meta.data,by="CellID")

SkinSeuratTotal@meta.data <- inner_join(x=skin.meta.data,y=meta.data,by="CellID") %>% column_to_rownames("CellID")
```

```{r}
#SUBSET DATA TO ONLY HEALTHY SAMPLES
travis_NM_skin_scRNA<- subset(SkinSeuratTotal,subset = Condition == c("Normal"))
```


```{r}
## PROCESS SC RNA DATA
travis_NM_skin_scRNA <- NormalizeData(travis_NM_skin_scRNA)
travis_NM_skin_scRNA <- FindVariableFeatures(travis_NM_skin_scRNA, selection.method = "vst", nfeatures = 2000)
travis_NM_skin_scRNA <- ScaleData(travis_NM_skin_scRNA)
travis_NM_skin_scRNA <- RunPCA(travis_NM_skin_scRNA, features = VariableFeatures(object = travis_NM_skin_scRNA))
travis_NM_skin_scRNA <- FindNeighbors(travis_NM_skin_scRNA, dims = 1:40)
travis_NM_skin_scRNA <- FindClusters(travis_NM_skin_scRNA, resolution = 1)
travis_NM_skin_scRNA <- RunUMAP(travis_NM_skin_scRNA, dims = 1:40)
```

## Normal skin - scRNA data from Travis paper 
```{r}
Idents(travis_NM_skin_scRNA) <- "Specific_CellType"
DimPlot(travis_NM_skin_scRNA,pt.size = 2.5,label = TRUE)

Idents(travis_NM_skin_scRNA) <- "CellType"
DimPlot(travis_NM_skin_scRNA,pt.size = 2.5)
```

```{r,eval=FALSE}

#saveRDS(travis_NM_skin_scRNA,file = "../../TRAVIS_scRNA_DATA/travis_NM_skin_scRNA.RDS")
```

```{r}
travis_NM_skin_scRNA <- readRDS(file="../../TRAVIS_scRNA_DATA/travis_NM_skin_scRNA.RDS")
```

```{r}
single_cell.markers <- FindAllMarkers(travis_NM_skin_scRNA,assay = "RNA",logfc.threshold = 0.25)
```

```{r}
filtered_single_cell.markers <- single_cell.markers %>% filter(p_val_adj<=0.05) %>% group_by(cluster) %>% top_n(n =300,wt = avg_log2FC) %>% filter(avg_log2FC>0.25)
filtered_spatial_markers <- new.skin.combined.markers %>% filter(p_val_adj<=0.05) %>% group_by(cluster) %>% top_n(n =300,wt = avg_log2FC) %>% filter(avg_log2FC>0.25)
```

# Figure - 2B
```{r}
##INTERSECT GENES BETWEEN scRNA and Spatial data
st.genes <- unique(rownames(new.skin.combined@assays$Spatial@counts))
sc.genes <- unique(rownames(travis_NM_skin_scRNA@assays$RNA@counts))
all.genes.scrna_and_spt <- unique(intersect(sc.genes,st.genes))
```
```{r}
MIA_results <- MIA(total_genes = length(all.genes.scrna_and_spt),single_cell.markers = filtered_single_cell.markers,spatial.markers = filtered_spatial_markers)

E.data <- MIA_results %>% column_to_rownames("cluster")
E.data <- E.data[,order(colnames(E.data))]
pheatmap(E.data,cluster_cols = FALSE,cluster_rows = FALSE,fontsize=15,color = col.pal)
```

```{r}
immune_only.E.data <- E.data[,c("T cell-1","T cell-2","T cell-3","Myeloid-1","Langerhans","Mast-1","Mast-2")]
pdf(file = "MIA_regions_IMMUNE_CELLS.pdf",width = 10,height = 4)
pheatmap(immune_only.E.data,cluster_cols = FALSE,cluster_rows = FALSE,fontsize=15,color = col.pal)
dev.off()
```

```{r}
structure_only.E_data <- E.data[,c("Fibroblast-1","Fibroblast-2","Fibroblast-4","Fibroblast-5","HairFollicle","Keratinocyte-2","Keratinocyte-3","Keratinocyte-5","Keratinocyte-6","Keratinocyte-8","Sebocyte","Venule-1","Venule-2","Venule-3","Venule-4","VSMC-1")]
pdf(file = "MIA_regions_STRUCTURAL_CELLS.pdf",width = 10,height = 4)
pheatmap(structure_only.E_data,cluster_cols = FALSE,cluster_rows = FALSE,fontsize=15,color = col.pal)
dev.off()
```

## NOW CALCULATING ENRICHMENT USING NEW DATASET (Travis et al)
# Seurat based cell type enrichment
```{r,eval=FALSE}
# RUN THIS ON BIG PURPLE 
skin_reference.2 <- SCTransform(travis_NM_skin_scRNA, ncells = 3000, verbose = FALSE) %>% RunPCA(verbose = FALSE) %>% 
    RunUMAP(dims = 1:40)
anchors <- FindTransferAnchors(reference = skin_reference.2, query = new.skin.combined, normalization.method = "SCT")
predictions.assay <- TransferData(anchorset = anchors, refdata = skin_reference.2$Specific_CellType, prediction.assay = TRUE, 
    weight.reduction = new.skin.combined[["pca"]],dims = 1:40)
new.skin.combined[["predictions_travis_data"]] <- predictions.assay
```

## CELL TYPES FOR FIGURE 1 - Travis et al. single cell dataset (Only using the healthy subset of scRNA data)
```{r}
Cell_types <- c("Fibroblast-1","HairFollicle","Keratinocyte-5","Keratinocyte-3","Sebocyte","VSMC-1","Melanocyte")
DefaultAssay(new.skin.combined) <- "predictions_travis_data"
for (x in Cell_types){
  pdf(file = paste("HEALTHY_FEMALE_1_R2_CELL_TYPE_",x,".pdf"),width = 10,height = 15)
  print(SpatialFeaturePlot(new.skin.combined, features = c(x), pt.size.factor = 2.5, ncol = 2, crop = TRUE,images = "HealthyFemale.1.S2") + scale_fill_gradientn(colors=cols,limits = c(0,1.01))) 
  dev.off()
}
```

# Cell counts for travis healthy subset data
```{r}
cell.counts <- as.data.frame(table(travis_NM_skin_scRNA@meta.data$Specific_CellType))
write.csv(cell.counts,file="../../TRAVIS_scRNA_DATA/CELL_COUNTS_TRAVIS_HEALTHY_SKIN_DATA.csv")
```

# Filtered out cell types with less than 50 cells
```{r}
filtered_cell.counts <- cell.counts %>% filter(Freq>50) 
Idents(travis_NM_skin_scRNA) <- "Specific_CellType"
travis_healthy_data.subset <- subset(x = travis_NM_skin_scRNA, idents = c(filtered_cell.counts$Var1))

DimPlot(travis_healthy_data.subset,pt.size = 2.5)
```

# Plotting final travis healthy data UMAP
# FIGURE -SUPPLEMENTARY S3
```{r}
pdf(width = 12,height=8,file = "UMAP_TRAVIS_DATA_HEALTHY_SAMPLES_ONLY.pdf")
DimPlot(travis_healthy_data.subset,pt.size = 3.5)
dev.off()
```

# Number of genes used in MIA - (Needed for Venn diagram)
```{r,eval=FALSE}

cluster.5 <- filtered_spatial_markers %>% filter(cluster==5)
Myeloid.1 <- filtered_single_cell.markers %>% filter(cluster=="Myeloid-1")

length(intersect(cluster.5$gene,Myeloid.1$gene))

cluster.6 <- filtered_spatial_markers %>% filter(cluster==6)
Keratinocyte.5 <- filtered_single_cell.markers %>% filter(cluster=="Keratinocyte-5")

length(intersect(cluster.6$gene,Keratinocyte.5$gene))
```

# SINGLE CELL DATA - ENRICHMENT PLOTS (USING ONLY ONE HEALTHY SAMPLE)
```{r}
HEALTHY.Female1.s2.clean<- st_filter_by_genes(st.data = HEALTHY.Female1.s2.clean,x = 200)
HEALTHY.Female1.s2.clean <- SCTransform(HEALTHY.Female1.s2.clean, verbose = FALSE,assay="Spatial")
HEALTHY.Female1.s2.clean <- RunPCA(HEALTHY.Female1.s2.clean)
```

```{r}
travis_NM_skin_scRNA <- readRDS(file="../../TRAVIS_scRNA_DATA/travis_NM_skin_scRNA.RDS")

# RUN THIS ON BIG PURPLE 
skin_reference.3 <- SCTransform(travis_NM_skin_scRNA, ncells = 3000, verbose = FALSE) %>% RunPCA(verbose = FALSE) %>% 
    RunUMAP(dims = 1:40)
anchors <- FindTransferAnchors(reference = skin_reference.3, query = HEALTHY.Female1.s2.clean, normalization.method = "SCT")
predictions.assay <- TransferData(anchorset = anchors, refdata = skin_reference.3$Specific_CellType, prediction.assay = TRUE, 
    weight.reduction = HEALTHY.Female1.s2.clean[["pca"]],dims = 1:40)
HEALTHY.Female1.s2.clean[["predictions_travis_data"]] <- predictions.assay

```

```{r}
SpatialColors <- colorRampPalette(colors = rev(x = brewer.pal(n = 11, name = "Spectral")))
cols <- SpatialColors(n = 100)
```

```{r}
SpatialFeaturePlot(HEALTHY.Female1.s2.clean,features = c("Melanocyte"), pt.size.factor = 2.5)+ scale_fill_gradientn(colors=cols,limits = c(0,1.000001))

SpatialFeaturePlot(HEALTHY.Female1.s2.clean,features = c("VSMC-1"), pt.size.factor = 2.5)+ scale_fill_gradientn(colors=cols,limits = c(0,1.000001))

SpatialFeaturePlot(HEALTHY.Female1.s2.clean,features = c("HairFollicle"), pt.size.factor = 2.5)+ scale_fill_gradientn(colors=cols,limits = c(0,1.000001))

SpatialFeaturePlot(HEALTHY.Female1.s2.clean,features = c("Keratinocyte-5"), pt.size.factor = 2.5)+ scale_fill_gradientn(colors=cols,limits = c(0,1.000001))

SpatialFeaturePlot(HEALTHY.Female1.s2.clean,features = c("Sebocyte"), pt.size.factor = 2.5)+ scale_fill_gradientn(colors=cols,limits = c(0,1.000001))

SpatialFeaturePlot(HEALTHY.Female1.s2.clean,features = c("Fibroblast-1"), pt.size.factor = 2.5)+ scale_fill_gradientn(colors=cols,limits = c(0,1.000001))
```

