---
title: "PS SAMPLES PART 5"
output: github_document
---

## PSORIASIS SAMPLES - CONTINUED - PART 5

## PSO + PSA + NORMAL SKIN COMBINED ANALYSIS

## Integration with PS skin scRNA data from (dataset 2) Reynolds et al.

## (DOI - <https://doi.org/10.1126/science.aba6500>).

| LINK / FIG NO | DESCRIPTION                                         |
|---------------|-----------------------------------------------------|
| [FIGURE S9C]  | UMAP for (PS only) dataset 2                        |
| [FIGURE 5B]   | MIA for (PS only) dataset 2 - Immune cell types     |
| [FIGURE S9E]  | MIA for (PS only) dataset 2 - Structural cell types |

### LOAD ALL PACKAGES

```{r}
library(tidyverse)
library(Seurat)
library(cowplot)
library(ggsci)
library(RColorBrewer)
library(pheatmap)
library(SeuratDisk)
```

### LOAD HELPER FUNCTIONS

```{r}
source("../SPATIAL_FUNCTIONS.R")
```

### LOAD COLOR PALETTE USED FOR ALL PLOTS

```{r}
col.pal <- RColorBrewer::brewer.pal(9, "OrRd")
```

### Integration with dataset 1 (Hughes et al)

Importing Harmonized (Batch corrected) ST data (produced in PS Samples part 1 notebook).

```{r}
# 1. All Spatial Samples with Harmony Batch Correction
skin_data.hm.sct <- readRDS(file = "/Volumes/Extreme Pro/GITHUB-DATA/ST-DATA/PSORIASIS-DATA/RDS-Files/ALL_SPATIAL_SAMPLES(HM_BATCH_CORRECTED).RDS")

# 2. Single Cell Markers
skin_data.hm.sct.markers <- readRDS("/Volumes/Extreme Pro/GITHUB-DATA/ST-DATA/PSORIASIS-DATA/RDS-Files/ALL_ST_HARMONY_ALIGNED_MARKERS.RDS")
```

```{r}
## COLOR FOR LABELS
color.labels <- c("0 Fibroblasts"="#87CEFA",
"1 Macs + fibroblasts"="#4876FF",
"2 Eccrine + melanocyte precursors"="#CD853F",
"3 Epidermis"="#BF96FF",
"4 Epidermis"="#FF0000",
"5 Connective tissue"="#CAF178",
"6 Mixed"="#E0BFB6",
"7 Epidermis"="#68228B",
"8 Hair follicle and sebaceous glands"="#7B0000",
"9 Adipose"="#FFC71A",
"10 Suprabasal keratinocytes"="#C355A0",
"11 Smooth muscle"="#00B923",
"12 Endothelial cells"="#8B5A2B",
"13 Immunoglobulins, fibroblasts"="#838B8B",
"14 Smooth muscle"="#005947",
"15 Mixed"="#C1CDCD",
"16 Adipose, fibroblasts"="#FF7545")
```

Load data-set 2 (Reynolds et al.) scRNA data

```{r}
library(SeuratDisk)
```

```{R,eval=FALSE}
hnf.data <- LoadH5Seurat("/Volumes/Extreme Pro/GITHUB-DATA/SC-RNA-DATA/HANIFFA-DATA/RDS-Files/submission.h5seurat")

hnf.data.PS <- subset(hnf.data,Status %in% c("Psoriasis"))
```

```{r,eval=FALSE}
hnf.data.PS <- NormalizeData(hnf.data.PS)
hnf.data.PS<- FindVariableFeatures(hnf.data.PS, selection.method = "vst", nfeatures = 2000)
hnf.data.PS<- ScaleData(hnf.data.PS)
hnf.data.PS<- RunPCA(hnf.data.PS, features = VariableFeatures(object = hnf.data.PS))
hnf.data.PS <- FindNeighbors(hnf.data.PS, dims = 1:40)
hnf.data.PS <- FindClusters(hnf.data.PS)
hnf.data.PS <- RunUMAP(hnf.data.PS, dims = 1:40)
```

```{r,eval=FALSE}
saveRDS(hnf.data.PS,file="HNF_SC_RNA_PSORIASIS_DATA.RDS")
```

```{r}
hnf.data.PS <- readRDS("/Volumes/Extreme Pro/GITHUB-DATA/SC-RNA-DATA/HANIFFA-DATA/RDS-Files/HNF_SC_RNA_PSORIASIS_DATA.RDS")
```

```{r}
hnf.data.PS.subset <- subset(hnf.data.PS,final_clustering != c("nan"))
```

### FIGURE S9C

```{r}
#pdf(width = 12,height=8,file = "UMAP_HANNIFA_DATA_PS_SAMPLES_ONLY.pdf")
DimPlot(hnf.data.PS.subset,group.by = "final_clustering",pt.size = 1.2,raster=FALSE)
#dev.off()
```

```{r}
DimPlot(hnf.data.PS.subset,group.by = "final_clustering",pt.size = 1.2,raster=FALSE)
```

```{r}
hnf.data.PS.subsampled <- hnf.data.PS.subset[, sample(colnames(hnf.data.PS.subset), size =25000, replace=F)]

#pdf(width = 12,height=8,file = "UMAP_HANNIFA_DATA_PS_SAMPLES_ONLY(AFTER_DOWNSAMPLING).pdf")
DimPlot(hnf.data.PS.subsampled,group.by = "final_clustering",pt.size = 3.5,raster=FALSE)
#dev.off()
```

```{r,eval=FALSE}
Idents(hnf.data.PS.subsampled) <- "final_clustering"
Marker_genes.HNF <- FindAllMarkers(hnf.data.PS.subsampled,max.cells.per.ident=1000,min.pct = 0.25)
saveRDS(Marker_genes.HNF,file="/Volumes/Extreme Pro/GITHUB-DATA/SC-RNA-DATA/HANIFFA-DATA/RDS-Files/Marker_genes.HNF.PS.RDS")
```

### MIA enrichment analysis with dataset 2

```{r}
Marker_genes.HNF <- readRDS(file = "/Volumes/Extreme Pro/GITHUB-DATA/SC-RNA-DATA/HANIFFA-DATA/RDS-Files/Marker_genes.HNF.PS.RDS")

filtered_single_cell.markers <- Marker_genes.HNF %>% filter(p_val_adj<=0.05) %>% group_by(cluster) %>% top_n(n =300,wt = avg_log2FC) %>% filter(avg_log2FC>0.25)
filtered_spatial_markers <- skin_data.hm.sct.markers %>% filter(p_val_adj<=0.05) %>% group_by(cluster) %>% top_n(n =300,wt = avg_log2FC) %>% filter(avg_log2FC>0.25)

##INTERSECT GENES BETWEEN scRNA and Spatial data
st.genes <- unique(rownames(skin_data.hm.sct@assays$Spatial@counts))
sc.genes <- unique(rownames(hnf.data.PS@assays$RNA@counts))
all.genes.scrna_and_spt <- unique(intersect(sc.genes,st.genes))

MIA_results <- MIA(total_genes = length(all.genes.scrna_and_spt),single_cell.markers = filtered_single_cell.markers,spatial.markers = filtered_spatial_markers)

E.data <- MIA_results %>% column_to_rownames("cluster")
E.data <- E.data[,order(colnames(E.data))]
pheatmap(E.data,cluster_cols = FALSE,cluster_rows = FALSE,fontsize=15,color = col.pal)
```

### FIGURE 5B

MIA (Immune cells only)

```{r}
immune_only.E.data <- E.data[,c("DC1","DC2","Macro_1","Macro_2","moDC_1","moDC_2","moDC_3","MigDC","Mono","Inf_mono","ILC1_NK","ILC2","ILC1_3","ILC2","Tc","Th","Treg","Mast_cell","NK")]
#pdf(file = "MIA_regions_IMMUNE_CELLS_HANIFFA_DATA.pdf",width = 10,height = 10)
pheatmap(immune_only.E.data,cluster_cols = FALSE,cluster_rows = FALSE,fontsize=15,color = col.pal)
#dev.off()
```

### FIGURE S9E

MIA (Structural cells only)

```{r}
structure_only.E_data <- E.data %>% dplyr::select(-c("DC1","DC2","Macro_1","Macro_2","moDC_1","moDC_2","moDC_3","MigDC","Mono","Inf_mono","ILC1_NK","ILC2","ILC1_3","ILC2","Tc","Th","Treg","Mast_cell","NK"))
#pdf(file = "MIA_regions_STRUCTURAL_CELLS_HANIFFA_DATA.pdf",width = 10,height = 10)
pheatmap(structure_only.E_data,cluster_cols = FALSE,cluster_rows = FALSE,fontsize=15,color = col.pal)
#dev.off()
```
