---
title: "R Notebook"
output: github_document
---

# PSO + PSA + NORMAL SKIN COMBINED ANALYSIS - PART 2

# CONTINUED

## Harmony-batch correction

## Table of content

| FIGURE NO | DESCRIPTION                                                             | LINK                       |
|----------|-----------------------------------|------------------------|
| S4A       | UMAP WITH ALL SPATIAL SAMPLES (Using Harmony batch correction)          | [FIGURE 3B and S4A]        |
| 3B        | UMAP WITH ALL SPATIAL SAMPLES (Using Harmony batch correction)          | [FIGURE 3B]                |
| 3A        | NON-LESIONAL SKIN SAMPLE (ST 21 NL), and LESIONAL SKIN SAMPLE (ST 22 L) | [FIGURE 3A ]               |
| S5        | Heatmap (After Harmony batch correction)                                | [SUPPLEMENTARY FIGURE S5 ] |
| 3C        | Percentage composition                                                  | [FIGURE 3C]                |
| S7B       | UMIs per cluster                                                        | [FIGURE S7B]               |

## LOAD ALL PACKAGES

```{r}
library(tidyverse)
library(Seurat)
library(cowplot)
library(ggsci)
library(ggpubr)
library(harmony)
```

## LOAD DATA (Generated in PS_SAMPLES_PART_1)

```{r,eval=FALSE}
All_Samples <- readRDS("/Volumes/Extreme Pro/GITHUB-DATA/ST-DATA/PSORIASIS-DATA/RDS-Files/All_Samples_list.RDS")
```

### Merging data for Harmony Batch correction

```{r,eval=FALSE}
all_skin.merge <- merge(x=All_Samples[[1]], y=All_Samples[-1], merge.data=TRUE)
```

### Save RDS File of All Merged Samples

```{r}
#saveRDS(all_skin.merge, file = "/gpfs/data/naiklab/Alex/GITHUB-DATA/ST-DATA/PSORIASIS-DATA/RDS-Files/all_samples_merged_obj.RDS")
```

### Load All Merged Samples & Run SCTransform

```{r}
all_skin.merge <- readRDS("/Volumes/Extreme Pro/GITHUB-DATA/ST-DATA/PSORIASIS-DATA/RDS-Files/all_samples_merged_obj.RDS")

all_skin.merge <- SCTransform(all_skin.merge,assay = "Spatial", vars.to.regress = "sample.id")
all_skin.merge <- RunPCA(all_skin.merge,assay = "SCT")
```

## Code for the final figure

### Running Harmony batch correction

```{r}
all_samples.hm.sct <- RunHarmony(all_skin.merge,assay.use = "SCT",project.dim = FALSE,group.by.vars = "sample.id")
```

```{r}
ElbowPlot(all_samples.hm.sct,ndims = 50)
```

```{r}
all_samples.hm.sct <- all_samples.hm.sct %>% 
    RunUMAP(reduction = "harmony", dims = 1:40) %>% 
    FindNeighbors(reduction = "harmony", dims = 1:40) %>% 
    FindClusters(resolution = 0.4) %>% 
    identity()
```

```{r}
DimPlot(all_samples.hm.sct)
```

```{r}
DimPlot(all_samples.hm.sct,split.by = "DISEASE_STATUS")
```

## IMPORT Pre-computed harmony results to get same embeddings and cluster labels as presented in the paper.

Pre-computed harmony batch results used in the final manuscript

```{r}
skin_data.hm.sct <- readRDS(file="/Volumes/Extreme Pro/GITHUB-DATA/ST-DATA/PSORIASIS-DATA/RDS-Files/ALL_SPATIAL_SAMPLES(HM_BATCH_CORRECTED).RDS")
```

```{r}
DimPlot(skin_data.hm.sct)
```

```{r}
## COLOR FOR LABELS
color.labels.heatmap <- c("#87CEFA",
"#4876FF",
"#CD853F",
"#BF96FF",
"#FF0000",
"#CAF178",
"#E0BFB6",
"#68228B",
"#7B0000",
"#FFC71A",
"#C355A0",
"#00B923",
"#8B5A2B",
"#838B8B",
"#005947",
"#C1CDCD",
"#FF7545")
```

## Find Markers for Each Cluster & Plot Heat Map of Top 10 Markers Per Cluster

```{r,fig.height=22,fig.width=15}
DefaultAssay(skin_data.hm.sct) <- "SCT"
Idents(skin_data.hm.sct) <- "seurat_clusters"
skin_data.hm.sct.markers <- FindAllMarkers(skin_data.hm.sct, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25,assay = "SCT") %>% filter(p_val_adj<0.05)

top10 <- skin_data.hm.sct.markers %>%
    group_by(cluster) %>%
    filter(gene %in% rownames(skin_data.hm.sct@assays$SCT@scale.data)) %>%
    top_n(n = 10, wt = avg_log2FC)

DoHeatmap(skin_data.hm.sct, features = top10$gene,assay = "SCT",group.colors = color.labels.heatmap,angle=90) + NoLegend()
```

### Defining cluster labels based on marker genes

```{r}
## ADDING CLUSTER LABELS
cluster.labels <- c("0 Fibroblasts",
"1 Macs + fibroblasts",
"2 Eccrine + melanocyte precursors",
"3 Epidermis",
"4 Epidermis",
"5 Connective tissue",
"6 Mixed",
"7 Epidermis",
"8 Hair follicle and sebaceous glands",
"9 Adipose",
"10 Suprabasal keratinocytes",
"11 Smooth muscle",
"12 Endothelial cells",
"13 Immunoglobulins, fibroblasts",
"14 Smooth muscle",
"15 Mixed",
"16 Adipose, fibroblasts")

## COLOR FOR LABELS
color.labels <- c("0 Fibroblasts"="#87CEFA",
"1 Macs + fibroblasts"="#4876FF",
"2 Eccrine + melanocyte precursors"="#CD853F",
"3 Epidermis"="#BF96FF",
"4 Epidermis"="#FF0000",
"5 Connective tissue"="#CAF178",
"6 Mixed"="#E0BFB6",
"7 Epidermis"="#68228B",
"8 Hair follicle and sebaceous glands"="#7B0000",
"9 Adipose"="#FFC71A",
"10 Suprabasal keratinocytes"="#C355A0",
"11 Smooth muscle"="#00B923",
"12 Endothelial cells"="#8B5A2B",
"13 Immunoglobulins, fibroblasts"="#838B8B",
"14 Smooth muscle"="#005947",
"15 Mixed"="#C1CDCD",
"16 Adipose, fibroblasts"="#FF7545")
```

```{r}
names(cluster.labels) <- levels(skin_data.hm.sct)
skin_data.hm.sct <- RenameIdents(skin_data.hm.sct, cluster.labels)
skin_data.hm.sct[["Spatial.regions"]] <- Idents(object = skin_data.hm.sct)
Idents(skin_data.hm.sct) <- "Spatial.regions"
```

```{r,fig.height=5,fig.width=10}
DimPlot(skin_data.hm.sct, split.by="DISEASE_STATUS",cols=color.labels, pt.size=3.5)
```

## FIGURE 3B and S4A

## FIGURE 3B {data-link="FIGURE 3B and S4A"}

### UMAP WITH ALL SPATIAL SAMPLES (AFTER HARMONY BATCH CORRECTION)

<a id="3b">

```{r,fig.height=7,fig.width=10}
DimPlot(skin_data.hm.sct, cols=color.labels, pt.size=3.5)
```

</a>

## FIGURE 3A 

### NON-LESIONAL SKIN SAMPLE (ST 21 NL)

<a id="3a">

```{r}
SpatialDimPlot(skin_data.hm.sct,images = c("ST_21_NL_Batch_6"),cols=color.labels,pt.size.factor = 2.5)
```

## FIGURE 3A

### LESIONAL SKIN SAMPLE (ST 22 L)

```{r}
SpatialDimPlot(skin_data.hm.sct,images = c("ST_22L_Batch_8"),cols=color.labels,pt.size.factor = 2.5)
```

</a>

### LESIONAL SKIN SAMPLE (ST 22 L)

```{r}
SpatialDimPlot(skin_data.hm.sct,images = c("ST_22L_Batch_8"), cols=color.labels,pt.size.factor = 2.5)
```

```{r}
#pdf(width = 8,height=10,file = "HEALTHY_FEMALE_SAMPLE_1_R2_SPATIAL_PLOT(HM_CORRECTED).pdf")
SpatialDimPlot(skin_data.hm.sct, images="ST.HF.1.R2", cols = color.labels,pt.size.factor = 2.4)
#dev.off()
```

### Get Top Markers for Harmony Batch Corrected Samples With Labeled Clusters

```{r}
DefaultAssay(skin_data.hm.sct) <- "SCT"
Idents(skin_data.hm.sct) <- "Spatial.regions"

top10 <- skin_data.hm.sct.markers %>%
    group_by(cluster) %>%
    filter(gene %in% rownames(skin_data.hm.sct@assays$SCT@scale.data)) %>%
    top_n(n = 10, wt = avg_log2FC)
```

## SUPPLEMENTARY FIGURE S5 

### HEATMAP showing top marker genes after Harmony batch correction

<a id="s5">

```{r,fig.height=20,fig.width=14}
DoHeatmap(skin_data.hm.sct, features = top10$gene,assay = "SCT", group.colors = color.labels,angle=90) + NoLegend()
```

</a>

### Fraction of spots per group (DISEASE STATUS)

```{r}
seurat_clusters.df.v2 <- table(skin_data.hm.sct@active.ident, skin_data.hm.sct@meta.data$DISEASE_STATUS,skin_data.hm.sct@meta.data$sample.id) %>% as.data.frame() %>%
  dplyr::rename(Cluster_id=Var1,Group=Var2,Sample_id=Var3) %>% filter(Freq!=0)  %>% group_by(Sample_id)%>% mutate(Fraction=Freq*100/sum(Freq)) %>% mutate(SUM_OF_FRACTIONS=sum(Fraction)) 
```

## FIGURE 3C

### PERCENTAGE COMPOSITION PLOT

<a id="3c">

```{r,fig.height=20,fig.width=7}
black.bold.16.text <- element_text(face = "bold", color = "black", size = 14,angle = 90, vjust = 0.5, hjust=1)
brks <- c(0, 0.25, 0.5, 0.75, 1)

ggplot(seurat_clusters.df.v2,aes(x=Group,y=Freq,fill=Cluster_id)) + geom_bar(stat="identity",position="fill") + ggplot2::theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(),axis.text.x =black.bold.16.text) + scale_y_continuous(breaks = brks, labels = scales::percent(brks)) + scale_fill_manual(values = color.labels)
```

</a>

### FIGURE S7B

### (1) UMIs per cluster

<a id="s7b">

### (2) Number of genes expressed per cluster

```{r fig.height=12,fig.width=12}
VlnPlot(skin_data.hm.sct,features =  "nFeature_Spatial") + scale_fill_manual(values = color.labels)

VlnPlot(skin_data.hm.sct,features =  "nCount_Spatial") + scale_fill_manual(values = color.labels)
```

</a>

```{r, eval=FALSE}
#saveRDS(skin_data.hm.sct.markers, "/gpfs/data/naiklab/Alex/GITHUB-DATA/ST-DATA/PSORIASIS-DATA/RDS-Files/ALL_ST_HARMONY_ALIGNED_MARKERS.RDS")
```

## Spatial plots with cluster labels

```{r}
Images(skin_data.hm.sct)
```

```{r}
## HEALTHY SAMPLE
SpatialDimPlot(skin_data.hm.sct, pt.size.factor = 2.5, crop = TRUE, images = "ST.HF.1.R2", cols = color.labels) 

## NON-LESIONAL PSO SAMPLE
# PSO_Non_Lesional_Patient5_Batch_6 (SAMPLE ID NEEDS TO BE CORRECTED)
SpatialDimPlot(skin_data.hm.sct,  pt.size.factor = 2.5, crop = TRUE, images = "ST_21_NL_Batch_6", cols = color.labels)

## LESIONAL PSO SAMPLE
# PSO_Lesional_Patient5_Batch_6 (SAMPLE ID NEEDS TO BE CORRECTED)
SpatialDimPlot(skin_data.hm.sct, pt.size.factor = 2.5, crop = TRUE, images = "ST_21_L_Batch_6", cols = color.labels)

## LESIONAL PSA SAMPLE
# PSA_LesionalSkinPatient1_Batch_1
SpatialDimPlot(skin_data.hm.sct, pt.size.factor = 2.5, crop = TRUE, images = "ST_16_L_Batch_1", cols = color.labels)

## NON-LESIONAL PSA SAMPLE
SpatialDimPlot(skin_data.hm.sct, pt.size.factor = 2.5, crop = TRUE, images = "ST_16_NL_Batch_1", cols = color.labels)
```

## ALTERNATIVE APPROACH

Not used in the final manuscript

```{r,eval=FALSE}
all_skin.list <-  lapply(X = All_Samples, 
                       FUN = SCTransform, 
                       method = "glmGamPoi", 
                       return.only.var.genes = FALSE,assay="Spatial",verbose=FALSE)

features <- SelectIntegrationFeatures(object.list = all_skin.list, nfeatures = 7500)
all_skin.merge <- merge(x=all_skin.list[[1]],y=all_skin.list[2:length(all_skin.list)], merge.data=TRUE)

VariableFeatures(all_skin.merge) <- features
all_skin.merge <- RunPCA(all_skin.merge,assay = "SCT")
all_samples.hm.sct <- RunHarmony(all_skin.merge,assay.use = "SCT",project.dim = FALSE, group.by.vars = "sample.id")
```

```{r,eval=FALSE}
all_samples.hm.sct <- all_samples.hm.sct %>% 
    RunUMAP(reduction = "harmony", dims = 1:40) %>% 
    FindNeighbors(reduction = "harmony", dims = 1:40) %>% 
    FindClusters(resolution = 0.5) %>% 
    identity()
```

```{r,eval=FALSE}
DimPlot(all_samples.hm.sct, split.by = "DISEASE_STATUS")
```

### Fraction of spots per group (DISEASE STATUS)

```{r,eval=FALSE}
seurat_clusters.df.v1 <- table(all_samples.hm.sct@active.ident, all_samples.hm.sct@meta.data$DISEASE_STATUS,all_samples.hm.sct@meta.data$sample.id) %>% 
  as.data.frame() %>%
  dplyr::rename(Cluster_id=Var1,Group=Var2,Sample_id=Var3) %>% 
  filter(Freq!=0) %>% 
  group_by(Sample_id) %>% 
  mutate(Fraction=Freq*100/sum(Freq)) %>% 
  mutate(SUM_OF_FRACTIONS=sum(Fraction)) 
```

```{r,fig.height=20,fig.width=7,eval=FALSE}
black.bold.16.text <- element_text(face = "bold", color = "black", size = 14,angle = 90, vjust = 0.5, hjust=1)
brks <- c(0, 0.25, 0.5, 0.75, 1)

ggplot(seurat_clusters.df.v1,aes(x=Group,y=Freq,fill=Cluster_id)) + geom_bar(stat="identity",position="fill") + ggplot2::theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(),axis.text.x =black.bold.16.text) + scale_y_continuous(breaks = brks, labels = scales::percent(brks))
```
