---
title: "PS SAMPLES PART 4"
output: github_document
---

# PS SAMPLES PART 4

## PSEUDO-BULK ANALYSIS

### *DIFFERENTIAL TESTING BETWEEN CONDITIONS - LESIONAL, NON-LESIONAL AND HEALTHY*

| FIGURE NO | DESCRIPTION                                                | LINK         |
|--------------|---------------------------------------------|--------------|
| 6A        | Hierarchical clustering based heatmap for all samples      | [link](#6a)  |
| 6B        | PCA plot for all samples - grouped by PS group             | [link](#6b)  |
| 6C        | PCA plot for all samples - grouped by Disease severity     | [link](#6c)  |
| 8A        | PCA plot for - Cluster 1 (Dermis macrophages, fibroblasts) | [link](#8a)  |
| 8B        | PCA plot for - Cluster 12 (Dermis lymphatics)              | [link](#8a)  |
| S9        | Dendogram for all samples                                  | [link](#s10) |
| S10       | PCA plots - per cluster basis (PS samples only)            | [link](#s10) |

### LOAD PACKAGES

```{r}
library(DESeq2)
library(Seurat)
library(tidyverse)
library(RColorBrewer)
library(pheatmap)
```

### LOAD HELPER FUNCTIONS

```{r}
source("~/Desktop/Ikjot/Helper_Files/SPATIAL_FUNCTIONS.R")
```

### LOAD DATA

```{r}
# 1. All Spatial Samples with Harmony Batch Correction
skin_data.hm.sct <- readRDS(file = "/gpfs/data/naiklab/Alex/GITHUB-DATA/ST-DATA/PSORIASIS-DATA/RDS-Files/ALL_SPATIAL_SAMPLES(HM_BATCH_CORRECTED).RDS")

# 2. All Spatial Samples Sample ID Meta Data
meta.data <- read.csv("/gpfs/data/naiklab/Alex/GITHUB-DATA/ST-DATA/PSORIASIS-DATA/Meta-Data/all_skin_corrected.meta.data.csv")
skin_data.hm.sct$sample.id <- meta.data$sample.id
```

### RE-NORMALIZING TO REGRESS OUT BATCH

```{r}
skin_data.hm.sct_re_normalized <- SCTransform(skin_data.hm.sct,assay = "Spatial", new.assay.name ="SCT_BATCH_REGRESSED", vars.to.regress = c("sample.id"))
```

### IDENTIFY DE GENES - LESIONAL VS HEALTHY

```{r,warning=FALSE,message=FALSE}
LES_vs_HEALTHY_v1 <- FindMarkers(skin_data.hm.sct_re_normalized, group.by = "DISEASE_STATUS", ident.1 = "Lesional", ident.2 = "Healthy skin")
LES_vs_HEALTHY_v2  <- FindMarkers(skin_data.hm.sct_re_normalized, group.by = "DISEASE_STATUS", ident.1 = "Lesional", ident.2 = "Healthy skin")

NON_LES_vs_HEALTHY <- FindMarkers(skin_data.hm.sct_re_normalized, group.by = "DISEASE_STATUS", ident.1 = "Non-Lesional", ident.2 = "Healthy skin")
LES_VS_NON_LES <- FindMarkers(skin_data.hm.sct_re_normalized, group.by = "DISEASE_STATUS", ident.1 = "Lesional", ident.2 = "Non-Lesional")
```

## PSEUDO-BULK APPROACH

### Get Pseudocounts & View Summary

```{r}
# Get Pseudocoounts groupin by sample.id
pseudo.counts <- Seurat:::PseudobulkExpression(object = skin_data.hm.sct, assays = "Spatial", group.by = "sample.id", slot = "counts", pb.method = "aggregate") 

# Pseudo Counts Data Frame
pseudo.counts.df <- as.data.frame(pseudo.counts$Spatial)

# View Summary of pseudocounts
summary(colSums(pseudo.counts$Spatial))
```

### Read & Process colData (Counts Meta Data)

```{r}
# Read PseudoBulk Meta Data, Filter Empty Sample IDs, & combine Group_1 & Severity into DISEASE_and_SEVERITY column
groups.table <- read.csv(file="/gpfs/data/naiklab/Alex/GITHUB-DATA/ST-DATA/PSORIASIS-DATA/Pseudo-Bulk-Data/groups.table.csv", stringsAsFactors = TRUE) %>% 
  filter(Sample.ID!="") %>%
  column_to_rownames("Sample.ID") %>%
  unite("DISEASE_and_SEVERITY", "GROUP_I", "SEVERITY", sep = "_", remove = FALSE, na.rm = FALSE)

# Add Sample IDs to be Sample.ID Column also
groups.table['Sample.ID'] <- rownames(groups.table)
```

### Match Counts Data Columns with colData rows

These need to be in the exact same order to create a DESeq2 Object

```{r}
# Match Samples (rownames) to be in same order as counts column names
groups.table <- groups.table[match(colnames(pseudo.counts.df), rownames(groups.table)),]
```

# DESeq Workflow

## 1. Create DESeq Object & Run DESeq2

```{r,,warning=FALSE,message=FALSE}
# 1. Create DESeq Object
dds <- DESeqDataSetFromMatrix(countData = pseudo.counts.df, colData=groups.table, design = ~ BATCH + GROUP_I)

# 2. Run DESeq2
dds <- DESeq(dds, quiet = TRUE)
```

## 2. Generate Transformed Counts for Downstream Analysis

```{r}
# 1. Get Normalized Counts
normalized.counts <- counts(dds, normalized=TRUE)
colSums(normalized.counts)

# 2. Variance Stabilized Transformation (VST) Object
dds.vst <- vst(dds, blind=TRUE)
```

### FIGURE 6B / 6C <a id="6b"></a><a id="6c"></a>

PCA plots showing clustering of samples based on disease group and severity / PASI score.

```{r,fig.height = 12,fig.width = 12}
# 1. PCA Plot grouping Skin Sample Type & Disease Status
plotPCA(dds.vst, intgroup=c("GROUP_I", "GROUP_II")) + 
  scale_color_manual(values=c("dimgray","#cc3333","#ffcc33","#ff9999","#ff9933"))

# 2. PCA Plot Grouped by DISEASE_and_SEVERITY scaled by PASI score
pcaData <- plotPCA(dds.vst, intgroup=c("SEVERITY", "PASI_SCORE", "DISEASE_and_SEVERITY", "Sample.ID"), returnData=TRUE)
percentVar <- round(100 * attr(pcaData, "percentVar"))

#pdf("ALL_SAMPLES_PCA_PLOT.pdf",height = 8,width = 10)
ggplot(pcaData, aes(PC1, PC2, fill = PASI_SCORE, shape = DISEASE_and_SEVERITY)) +
  geom_point(size=3) +
  xlab(paste0("PC1: ", percentVar[1], "% variance")) +
  ylab(paste0("PC2: ", percentVar[2], "% variance")) + 
  coord_fixed() +
  scale_shape_manual(values=c(21,22,23,24,25))
#dev.off()
```

```{r,fig.height = 12,fig.width = 12}
# 3. PCA Plot by DISEASE_and_SEVERITY Labeled Samples
#pdf("ALL_SAMPLES_PCA_PLOT(WITH_SAMPLE_IDS).pdf",height = 15,width = 25 )
ggplot(pcaData, aes(PC1, PC2,label=Sample.ID)) +
  geom_point(size=3,mapping = aes(color=DISEASE_and_SEVERITY)) +
  xlab(paste0("PC1: ",percentVar[1],"% variance")) +
  ylab(paste0("PC2: ",percentVar[2],"% variance")) + 
  coord_fixed() + geom_text(hjust=0.5, vjust=0) + 
  scale_color_manual(values=c("dimgray","#cc3333","#ffcc33","#ff9999","#ff9933"))
#dev.off()
```

### FIGURE 6A <a id="6a"></a>

### Hierarchical Clustering - Samples cluster by disease severity not systemic co-morbidity

```{r,fig.height = 20,fig.width = 12}
sampleDists <- dist(t(assay(dds.vst)))
sampleDistMatrix <- as.matrix(sampleDists)

rownames(sampleDistMatrix) <- paste(dds.vst$GROUP_I, dds.vst$SEVERITY,dds.vst$GROUP_II, sep="-")
colnames(sampleDistMatrix) <- NULL

colors <- colorRampPalette(brewer.pal(9, "RdYlBu")) (255)

#pdf("HC_WITH_HEATMAP.pdf",height = 15,width = 15)
print(pheatmap(sampleDistMatrix,
         clustering_distance_rows=sampleDists,
         clustering_distance_cols=sampleDists,
         col=colors))
#dev.off()
```

### FIGURE S9 <a id="s9"></a>

## DENDOGRAM

```{r,fig.height = 25,fig.width = 15}
hc <- hclust(sampleDists)

## Printing dendogram with - Disease Severity
print(plot(hc, labels=paste(dds.vst$GROUP_I, dds.vst$SEVERITY, dds.vst$GROUP_II,dds.vst$PASI_SCORE, sep = ":")))

## Printing Dendogram - with sample ids
#pdf("DENDOGRAM-WITH-SAMPLE-IDS.pdf",height = 14,width=10)
plot(hc, labels=paste(dds.vst$GROUP_I, dds.vst$Sample.ID, dds.vst$GROUP_II,dds.vst$PASI_SCORE, sep = ":"))
#dev.off() 
```

### Figure S9 <a id="s9"></a>

### CLUSTER SPECIFIC (PSEUDO-BULK) PCA-plots

#### PART 1

```{r}
# 1. All Cluster Labels
cluster.labels <- unique(skin_data.hm.sct@meta.data$Spatial.regions) %>% 
  as.vector() %>% 
  sort(decreasing = TRUE)

# 2. Cluster Labels for Specific Groups of Interest
cluster.labels.filtered <- c("9 Adipose","10 Suprabasal keratinocytes","11 Smooth muscle","12 Endothelial cells",
                             "13 Immunoglobulins, fibroblasts","14 Smooth muscle","15 Mixed","16 Adipose, fibroblasts")
```

```{r,warning=FALSE,message=FALSE,fig.height = 12,fig.width = 12}
# 1. File Path for Groups Table CSV
groups.tbl.path <- "/gpfs/data/naiklab/Alex/GITHUB-DATA/ST-DATA/PSORIASIS-DATA/Pseudo-Bulk-Data/groups.table.csv"

for(x in cluster.labels){
  subset.data <- subset(skin_data.hm.sct,Spatial.regions %in% c(x))
  dds <- pseudo_bulk_out(subset.data, group_label = "sample.id", groups_tbl_path = groups.tbl.path)
  if(!is.null(dds)){
    tryCatch({
      dds.vst <- vst(dds, blind=TRUE)
      
      # ALL PSORIASIS SAMPLES
      dds.vst_subset <- dds.vst[ ,dds.vst$GROUP_I %in% c("Lesional","Non-Lesional")]
      
      ## PCA PLOT
      #pdf(file=paste("PSEUDO_BULK_OUTPUT/ALL_PS_SAMPLES/",x,"_PCA_PLOT_PSEUDOBULK_ALL_SAMPLES.pdf"),height = 8,width = 10)
      print(plotPCA(dds.vst_subset, intgroup=c("GROUP_I", "SEVERITY"))+ ggtitle(paste("PCA Plot for -",x)) + 
              scale_color_manual(values=c("#ff9999","#cc3333","#ffcc33","#ff9933")))
      #dev.off()
      
    }, error=function(e){ skip_to_next <<- TRUE})
  }
}
```

#### PART-2

```{r,warning=FALSE,message=FALSE,fig.height = 12,fig.width = 12}
for(x in cluster.labels.filtered) {
  subset.data <- subset(skin_data.hm.sct,Spatial.regions %in% c(x))
  dds <- pseudo_bulk_out(subset.data,group_label = "sample.id", groups_tbl_path = groups.tbl.path)
  if(!is.null(dds)){
    tryCatch({
      dds.vst <- vst(dds, blind=TRUE,nsub = 100)
      normalized.counts <- counts(dds, normalized=TRUE)
      
      # ONLY PSORIASIS SAMPLES
      dds.vst_subset <- dds.vst[ ,dds.vst$GROUP_I %in% c("Lesional","Non-Lesional")]
      
      ## PCA PLOT
      #pdf(file=paste("PSEUDO_BULK_OUTPUT/ALL_SAMPLES(VERSION_2)/",x,"_PCA_PLOT_PSEUDOBULK_ALL_SAMPLES.pdf"),height = 8,width = 10)
      print(plotPCA(dds.vst_subset, intgroup=c("GROUP_I", "SEVERITY"))+ ggtitle(paste("PCA Plot for -",x)) +
              scale_color_manual(values=c("#ff9999","#cc3333","#ffcc33","#ff9933"))) 
      #dev.off()
      
    }, error=function(e){ skip_to_next <<- TRUE})
  }
}
```

### FIGURE 8A and 8B<a id="8a"></a>

```{r,warning=FALSE,message=FALSE,fig.height = 12,fig.width = 12}
subset.data <- subset(skin_data.hm.sct,Spatial.regions %in% c("1 Macs + fibroblasts"))

dds <- pseudo_bulk_out(subset.data,group_label = "sample.id", groups_tbl_path = groups.tbl.path)

dds.vst <- vst(dds, blind=TRUE, nsub = 100)

# ALL PSORIASIS SAMPLES
dds.vst_subset_2 <- dds.vst[,dds.vst$GROUP_I %in% c("Lesional","Non-Lesional")]
#pdf(file=paste("PSEUDO_BULK_OUTPUT/ALL_PS_SAMPLES/","1 Macs + fibroblasts","_PCA_PLOT_PSEUDOBULK_ALL_SAMPLES(WITH_LABELS)(V2).pdf"),height = 8,width = 10)
plotPCA(dds.vst_subset_2, intgroup=c("GROUP_I", "SEVERITY")) + geom_text(label = dds.vst_subset_2$PASI_SCORE,nudge_x = 0.50, nudge_y = 1) + scale_color_manual(values=c("#ff9999","#cc3333","#ffcc33","#ff9933"))
#dev.off()

#12 Endothelial cells
subset.data <- subset(skin_data.hm.sct,Spatial.regions %in% c("12 Endothelial cells"))
dds <- pseudo_bulk_out(subset.data,group_label = "sample.id", groups_tbl_path = groups.tbl.path)
dds.vst <- vst(dds, blind=TRUE,nsub = 100)

# ALL PSORIASIS SAMPLES
dds.vst_subset_2 <- dds.vst[ ,dds.vst$GROUP_I %in% c("Lesional","Non-Lesional")]

#pdf(file=paste("PSEUDO_BULK_OUTPUT/ALL_PS_SAMPLES/","12 Endothelial cells","_PCA_PLOT_PSEUDOBULK_ALL_SAMPLES(WITH_LABELS)(V2).pdf"),height = 8,width = 10)
plotPCA(dds.vst_subset_2, intgroup=c("GROUP_I", "SEVERITY")) + geom_text(label = dds.vst_subset_2$PASI_SCORE,nudge_x = 0.50, nudge_y = 1)  + scale_color_manual(values=c("#ff9999","#cc3333","#ffcc33","#ff9933"))
#dev.off()
```
