---
title: "TRAVIS HUGHES et al single cell- healthy skin subset"
output: github_document
---

# ST HEALTHY SAMPLES - PART 2

## Integration with TRAVIS HUGHES et al.

## (DOI - <https://doi.org/10.1016/j.immuni.2020.09.015>).

### Load all of the required packages

| FIGURE NO | LINK         |
|-----------|--------------|
| 2B        | [link](#2b)  |
| 3A        | [link](#3a)  |
| S3A       | [link](#s3a) |
| 1E        | [link](#1e)  |

```{r}
library(tidyverse)
library(Seurat)
library(cowplot)
library(RColorBrewer)
library(pheatmap)
library(scales)
```

### Import functions from the custom R script

```{r}
source("../SPATIAL_FUNCTIONS.R")
```

### Load Color Palette Used for Downstream Plots

```{r}
col.pal <- RColorBrewer::brewer.pal(9, "OrRd")
```

### Load Seurat object with healthy samples.

(*replace this with zenodo repo link for final release)*

```{r}
#new.skin.combined <- readRDS("/Volumes/Extreme Pro/GITHUB-DATA/ST-DATA/HEALTHY-DATA/RDS-Files/HEALTHY_SKIN_SAMPLES_ST_CLUSTERS_LABELED.RDS")
new.skin.combined <- readRDS("/Volumes/Extreme Pro/GITHUB-DATA/ST-DATA/HEALTHY-DATA/RDS-Files/HEALTHY_SKIN_SAMPLES_ST.RDS")
```

### Load Seurat object + meta-data for scRNA data from Travis et al

(*replace this with zenodo repo link for final release)*

```{r,eval=FALSE}
# Load Travis Data into Seurat Object
travis.sc.seu.obj <- readRDS("/Volumes/Extreme Pro/GITHUB-DATA/SC-RNA-DATA/TRAVIS-DATA/RDS-Files/Travis_Data_Seurat_Object.RDS")
```

```{r,eval=FALSE}
# META DATA FROM TRAVIS Paper 
meta.data <- read.csv("/Volumes/Extreme Pro/GITHUB-DATA/SC-RNA-DATA/TRAVIS-DATA/Meta-Data/Cell_Level_Metadata.csv")
```

```{r,eval=FALSE}
# Gather Travis Meta Data & Subset Psoriasis Data
skin.meta.data <- as.data.frame(travis.sc.seu.obj@meta.data) %>% rownames_to_column("CellID")
travis.sc.seu.obj@meta.data <- inner_join(x=skin.meta.data,y=meta.data,by="CellID") %>% column_to_rownames("CellID")

# Subset Data to have only Healthy Samples
travis_NM_skin_scRNA <- subset(travis.sc.seu.obj,subset = Condition == c("Normal"))
```

```{r,eval=FALSE}
## PROCESS SC RNA DATA
travis_NM_skin_scRNA <- NormalizeData(travis_NM_skin_scRNA)
travis_NM_skin_scRNA <- FindVariableFeatures(travis_NM_skin_scRNA, selection.method = "vst", nfeatures = 2000)
travis_NM_skin_scRNA <- ScaleData(travis_NM_skin_scRNA)
travis_NM_skin_scRNA <- RunPCA(travis_NM_skin_scRNA, features = VariableFeatures(object = travis_NM_skin_scRNA))
travis_NM_skin_scRNA <- FindNeighbors(travis_NM_skin_scRNA, dims = 1:40)
travis_NM_skin_scRNA <- FindClusters(travis_NM_skin_scRNA, resolution = 1)
travis_NM_skin_scRNA <- RunUMAP(travis_NM_skin_scRNA, dims = 1:40)
```

Save the processed Seurat object to an RDS file for quick import.

```{r,eval=FALSE}
#saveRDS(travis_NM_skin_scRNA, file = "/Volumes/Extreme Pro/GITHUB-DATA/SC-RNA-DATA/TRAVIS-DATA/RDS-Files/travis_NM_skin_scRNA.RDS")
```

Import the processed Seurat object

```{r}
travis_NM_skin_scRNA <- readRDS("/Volumes/Extreme Pro/GITHUB-DATA/SC-RNA-DATA/TRAVIS-DATA/RDS-Files/travis_NM_skin_scRNA.RDS")
```

Plot UMAP for normal scRNA data

```{r}
# Plot UMAP for Specific Cell Types Column in Meta Data
Idents(travis_NM_skin_scRNA) <- "Specific_CellType"
DimPlot(travis_NM_skin_scRNA,pt.size = 2.5,label = TRUE)

# Plot UMAP for Cell Types Column in Meta Data
Idents(travis_NM_skin_scRNA) <- "CellType"
DimPlot(travis_NM_skin_scRNA,pt.size = 2.5,label = TRUE)
```

### IDENTIFY MARKER GENES FOR ST DATA AND THE SC RNA DATA

```{r}
# 1. Load Healthy Samples Spatial Markers Found in ST_HEALTHY_SAMPLES_PART_1
filtered_spatial_markers <- readRDS("/Volumes/Extreme Pro/GITHUB-DATA/ST-DATA/HEALTHY-DATA/RDS-Files/HEALTHY_ST_MARKERS.RDS")
```

```{r,eval=FALSE}
# 2. Find All Markers from Travis Healthy Samples
Idents(travis_NM_skin_scRNA) <- "Specific_CellType"
single_cell.markers <- FindAllMarkers(travis_NM_skin_scRNA, assay = "RNA", logfc.threshold = 0.25)
#saveRDS(single_cell.markers, file="../../TRAVIS_scRNA_DATA/Travis_NM_skin_scRNA_markers.RDS")
```

```{r}
# 3. Load Saved Single Cell Healthy Markers
single_cell.markers <- readRDS("/Volumes/Extreme Pro/GITHUB-DATA/SC-RNA-DATA/TRAVIS-DATA/RDS-Files/Travis_NM_skin_scRNA_markers.RDS")
```

```{r}
# 4. Filter Single Cell Markers
filtered_single_cell.markers <- single_cell.markers %>% 
  filter(p_val_adj <= 0.05) %>% 
  group_by(cluster) %>% 
  top_n(n = 300, wt = avg_log2FC) %>% 
  filter(avg_log2FC > 0.25)
```

## Figure - 2B

### MIA ( Multi-modal intersection analysis)

#### original publication- (<https://doi.org/10.1038/s41587-019-0392-8>)

#### pre-computation parameters

#### Number of intersecting / common genes between ST and scRNA data -set as background genes.

```{r}
##INTERSECT GENES BETWEEN scRNA and Spatial data
st.genes <- unique(rownames(new.skin.combined@assays$Spatial@counts))
sc.genes <- unique(rownames(travis_NM_skin_scRNA@assays$RNA@counts))
all.genes.scrna_and_spt <- unique(intersect(sc.genes, st.genes))
```

<a id="2b">

```{r}
MIA_results <- MIA(total_genes = length(all.genes.scrna_and_spt), single_cell.markers = filtered_single_cell.markers,spatial.markers = filtered_spatial_markers)

E.data <- MIA_results %>% column_to_rownames("cluster")
E.data <- E.data[,order(colnames(E.data))]
pheatmap(E.data, cluster_cols = FALSE, cluster_rows = FALSE, fontsize=15, color = col.pal)
```

</a>

## Figure - 3A

### Immune cell types

<a id="3a">

```{r}
immune_only.E.data <- E.data[,c("T cell-1","T cell-2","T cell-3","Myeloid-1","Langerhans","Mast-1","Mast-2")]

pdf(file = "MIA_regions_IMMUNE_CELLS(HEALTHY-SAMPLES_ONLY).pdf",width = 10,height = 4)
pheatmap(immune_only.E.data,cluster_cols = FALSE,cluster_rows = FALSE,fontsize=15,color = col.pal)
dev.off()
```

</a>

### Strucutural cell types

```{r}
structure_only.E_data <- E.data[,c("Fibroblast-1","Fibroblast-2","Fibroblast-4","Fibroblast-5","HairFollicle","Keratinocyte-2","Keratinocyte-3","Keratinocyte-5","Keratinocyte-6","Keratinocyte-8","Sebocyte","Venule-1","Venule-2","Venule-3","Venule-4","VSMC-1")]

pdf(file = "MIA_regions_STRUCTURAL_CELLS(HEALTHY-SAMPLES_ONLY).pdf",width = 10,height = 4)
pheatmap(structure_only.E_data,cluster_cols = FALSE,cluster_rows = FALSE,fontsize=15,color = col.pal)
dev.off()
```

### Calculate enrichment for each cell type (Using Seurat Anchor Integration)

```{r,eval=FALSE}
# RUN THIS ON BIG PURPLE 
skin_reference.2 <- SCTransform(travis_NM_skin_scRNA, ncells = 3000, verbose = FALSE) %>% 
  RunPCA(verbose = FALSE) %>% 
  RunUMAP(dims = 1:40)

anchors <- FindTransferAnchors(reference = skin_reference.2, query = new.skin.combined, normalization.method = "SCT")
predictions.assay <- TransferData(anchorset = anchors, refdata = skin_reference.2$Specific_CellType, prediction.assay = TRUE, 
                                  weight.reduction = new.skin.combined[["pca"]], dims = 1:40)

new.skin.combined[["predictions_travis_data"]] <- predictions.assay
```

Color scale for cell type enrichment plots

```{r}
SpatialColors <- colorRampPalette(colors = rev(x = brewer.pal(n = 11, name = "Spectral")))
cols <- SpatialColors(n = 100)
```

## Figure -1E

<a id="1e">

```{r}
Cell_types <- c("Fibroblast-1","HairFollicle","Keratinocyte-5","Keratinocyte-3","Sebocyte","VSMC-1","Melanocyte")
DefaultAssay(new.skin.combined) <- "predictions_travis_data"

for (x in Cell_types){
  #pdf(file = paste("HEALTHY_FEMALE_1_R2_CELL_TYPE_",x,".pdf"),width = 10,height = 15)
  print(SpatialFeaturePlot(new.skin.combined, features = c(x), pt.size.factor = 2.5, ncol = 2, crop = TRUE, images = "HV2.S1.R2") +
    scale_fill_gradientn(colors=cols,limits = c(0, 1.01))) 
  #dev.off()
}
```

</a>

### Overall cell frequencies in healthy subset data

```{r}
cell.counts <- as.data.frame(table(travis_NM_skin_scRNA@meta.data$Specific_CellType))
#write.csv(cell.counts,"CELL_COUNTS_TRAVIS_HEALTHY_SKIN_DATA.csv")
```

### Filtered out cell types with less than 50 cells

```{r}
# Filter Healthy Subset by Cell Types with 50 or mroe cells
filtered_cell.counts <- cell.counts %>% filter(Freq > 50) 
Idents(travis_NM_skin_scRNA) <- "Specific_CellType"
travis_healthy_data.subset <- subset(x = travis_NM_skin_scRNA, idents = c(filtered_cell.counts$Var1))

# Plot UMAP
DimPlot(travis_healthy_data.subset, pt.size = 2.5)
```

### Supplementary Figure - S3A

<a id="s3a">

```{r}
#pdf(width = 12,height=8,file = "UMAP_TRAVIS_DATA_HEALTHY_SAMPLES_ONLY.pdf")
DimPlot(travis_healthy_data.subset, pt.size = 3.5, group.by = "CellType")
#dev.off()
```

</a>

### Data for Venn diagram - Figure 1F

```{r}
# Venn Diagram Data Intersecting Spatial Cluster 5 with Myeloid-1 Cluster
cluster.5 <- filtered_spatial_markers %>% filter(cluster==5)
Myeloid.1 <- filtered_single_cell.markers %>% filter(cluster=="Myeloid-1")

length(intersect(cluster.5$gene,Myeloid.1$gene))

# Venn Diagram Data Intersecting Spatial Cluster 6 with Keratinocyte-5 Cluster
cluster.6 <- filtered_spatial_markers %>% filter(cluster==6)
Keratinocyte.5 <- filtered_single_cell.markers %>% filter(cluster=="Keratinocyte-5")

length(intersect(cluster.6$gene,Keratinocyte.5$gene))
```
